---
output:
  html_document: default
  pdf_document: default
---
### Test Preliminary analysis

---
  title: "Single Nucleus Sequencing Trial Quality Control and Analysis"
output:
  html_document: default
github_document: default
---
  
```{r include = FALSE}
### Load data and packages

knitr::opts_chunk$set(fig.width=12, fig.height=8) 

```


```{r, include = FALSE, message = FALSE, warning = FALSE, cache = FALSE}
library(Seurat)
require(gplots)
require(ComplexHeatmap)
```

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE}

setwd('/home/jovyan/MH_DDD/')

tab = read.delim('../data/KptnMouse/RNAscope/IntensityInNuclei/Objects_Population - Nuclei.txt', sep = '\t', skip = 9)
tab = tab[order(-tab$Position.Y..µm.),]
```

Remove too large (doublets) and  too small (partial) nuclei:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE}

hist(tab$Nuclei...Nucleus.Volume..µm.., breaks = 100, main = 'Nuclei Volume histogram before filtering')
plot(tab$Position.Y..µm., tab$Position.X..µm., pch = '.', main = 'Nuclei position before filtering')
keep = tab$Nuclei...Nucleus.Volume..µm.. < 4000
print(sum(keep)/dim(tab)[1])
tab = tab[keep,]
keep = tab$Nuclei...Nucleus.Volume..µm.. > 0
print(sum(keep)/dim(tab)[1])
tab = tab[keep,]
hist(tab$Nuclei...Nucleus.Volume..µm.., breaks = 100, main = 'Nuclei Volume histogram after filtering')
plot(tab$Position.Y..µm., tab$Position.X..µm., pch = '.', main = 'Nuclei position after filtering')

```

Make 1D histogram plots of intensity (sum, mean or normalized sum) within nuclei for one field in the first section only:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE}

sectionNumber = rep(0, dim(tab)[1])
count = 1
sectionNumber[1] = count
for (i in 2:dim(tab)){
if (abs(tab$Position.Y..µm.[i] - tab$Position.Y..µm.[i-1]) > 1000){count = count + 1}
sectionNumber[i] = count
}

section = 1
print(table(tab$Field))
field = c(50)

meanIntensity_singleField = as.matrix(tab[fieldNumber %in% field,c('Nuclei...Intensity.Nucleus.Atto.490LS.Mean', 'Nuclei...Intensity.Nucleus.Atto.425.Mean',
                         'Nuclei...Intensity.Nucleus.Alexa.488.Mean', 'Nuclei...Intensity.Nucleus.Alexa.568.Mean',
                         'Nuclei...Intensity.Nucleus.Alexa.647.Mean')])
sumIntensity_singleField = as.matrix(tab[fieldNumber %in% field,c('Nuclei...Intensity.Nucleus.Atto.490LS.Sum', 'Nuclei...Intensity.Nucleus.Atto.425.Sum',
                         'Nuclei...Intensity.Nucleus.Alexa.488.Sum', 'Nuclei...Intensity.Nucleus.Alexa.568.Sum',
                         'Nuclei...Intensity.Nucleus.Alexa.647.Sum')])
normSumIntensity_singleField = sumIntensity_singleField/rowSums(sumIntensity_singleField)

names(meanIntensity_singleField) = names(sumIntensity_singleField) = names(normSumIntensity_singleField) = tab$Object.No[fieldNumber == field]

channels = c('490LS', '425', '488', '568', '647')
for (i in 1:length(channels)){
par(mfrow = c(1,3))
hist(meanIntensity_singleField[, i], 25, main = paste('Histogram of', channels[i], 'Mean Log-Intensity'))
hist(sumIntensity_singleField[, i], 25, main = paste('Histogram of', channels[i], 'Log-Intensity Sum'))
hist(normSumIntensity_singleField[, i], 25, main = paste('Histogram of', channels[i], 'Normalized Intensity Sum'))
}

channels = c('490LS', '425', '488', '568', '647')
for (i in 1:length(channels)){
par(mfrow = c(1,3))
plot(1:length(meanIntensity_singleField[,i]), meanIntensity_singleField[, i], main = paste('Histogram of', channels[i], 'Mean Log-Intensity'), pch = '.')
plot(1:length(sumIntensity_singleField[,i]), sumIntensity_singleField[, i], main = paste('Histogram of', channels[i], 'Log-Intensity Sum'), pch = '.')
plot(1:length(normSumIntensity_singleField[, i]), normSumIntensity_singleField[,i], main = paste('Histogram of', channels[i], 'Normalized Intensity Sum'), pch = '.')
}

par(mfrow = c(1,5))
for (i in 1:length(channels)){
plot(1:length(meanIntensity_singleField[,i]), meanIntensity_singleField[, i], main = paste(channels[i]), pch = '.')
}

channels = c('490LS', '425', '488', '568', '647')
for (i in 1:length(channels)){
par(mfrow = c(1,3))
hist(log(meanIntensity_singleField[, i],2), 25, main = paste('Histogram of', channels[i], 'Mean Log-Intensity'))
hist(log(sumIntensity_singleField[, i],2), 25, main = paste('Histogram of', channels[i], 'Log-Intensity Sum'))
hist(log(normSumIntensity_singleField[, i],2), 25, main = paste('Histogram of', channels[i], 'Normalized Intensity Sum'))
}

channels = c('490LS', '425', '488', '568', '647')
for (i in 1:length(channels)){
par(mfrow = c(1,3))
plot(1:length(meanIntensity_singleField[,i]), log(meanIntensity_singleField[, i],2), main = paste('Histogram of', channels[i], 'Mean Log-Intensity'), pch = '.')
plot(1:length(sumIntensity_singleField[,i]), log(sumIntensity_singleField[, i],2), main = paste('Histogram of', channels[i], 'Log-Intensity Sum'), pch = '.')
plot(1:length(normSumIntensity_singleField[, i]), log(normSumIntensity_singleField[,i],2), main = paste('Histogram of', channels[i], 'Normalized Intensity Sum'), pch = '.')
}

par(mfrow = c(1,5))
for (i in 1:length(channels)){
plot(1:length(meanIntensity_singleField[,i]), log(meanIntensity_singleField[, i],2), main = paste(channels[i]), pch = '.')
}

```

Make 2D histogram for the whole section:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE}

fieldNumber = tab$Field
print(table(fieldNumber))

meanIntensity = as.matrix(tab[sectionNumber == section,c('Nuclei...Intensity.Nucleus.Atto.490LS.Mean', 'Nuclei...Intensity.Nucleus.Atto.425.Mean',
                         'Nuclei...Intensity.Nucleus.Alexa.488.Mean', 'Nuclei...Intensity.Nucleus.Alexa.568.Mean',
                         'Nuclei...Intensity.Nucleus.Alexa.647.Mean')])
sumIntensity = as.matrix(tab[sectionNumber == section,c('Nuclei...Intensity.Nucleus.Atto.490LS.Sum', 'Nuclei...Intensity.Nucleus.Atto.425.Sum',
                         'Nuclei...Intensity.Nucleus.Alexa.488.Sum', 'Nuclei...Intensity.Nucleus.Alexa.568.Sum',
                         'Nuclei...Intensity.Nucleus.Alexa.647.Sum')])
normSumIntensity = sumIntensity/rowSums(sumIntensity)


names(meanIntensity) = names(sumIntensity) = names(normSumIntensity) = tab$Field[sectionNumber == section]

channels = c('490LS', '425', '488', '568', '647')
for (i in 1:length(channels)){
par(mfrow = c(1,3))
hist(meanIntensity[, i], 100, main = paste('Histogram of', channels[i], 'Mean Log-Intensity'))
hist(sumIntensity[, i], 100, main = paste('Histogram of', channels[i], 'Log-Intensity Sum'))
hist(normSumIntensity[, i], 100, main = paste('Histogram of', channels[i], 'Normalized Intensity Sum'))
}

for (i in 1:length(channels)){
par(mfrow = c(1,3))
plot(1:length(meanIntensity[,i]), meanIntensity[, i], main = paste('Histogram of', channels[i], 'Mean Log-Intensity'), pch = '.')
plot(1:length(sumIntensity[,i]), sumIntensity[, i], main = paste('Histogram of', channels[i], 'Log-Intensity Sum'), pch = '.')
plot(1:length(normSumIntensity[, i]), normSumIntensity[,i], main = paste('Histogram of', channels[i], 'Normalized Intensity Sum'), pch = '.')
}

par(mfrow = c(1,5))
for (i in 1:length(channels)){
plot(1:length(meanIntensity[,i]), meanIntensity[, i], main = paste(channels[i]), pch = '.')
}

```

Now make 2D histograms and scatter plots for each combination of channels:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE}

par(mfrow = c(5,5), mar=c(2,2,2,2))
for (i in 1:dim(meanIntensity)[2]){
  for (j in 1:dim(meanIntensity)[2]){
    hist2d(meanIntensity[, c(i,j)], xlab = channels[i], ylab = channels[j], bins = 50,
           main = paste(channels[i], 'vs.', channels[j]))
  }
}

par(mfrow = c(5,5), mar=c(2,2,2,2))
for (i in 1:dim(meanIntensity)[2]){
  for (j in 1:dim(meanIntensity)[2]){
    plot(meanIntensity[, i], meanIntensity[,j], xlab = channels[i], ylab = channels[j], bins = 50,
           main = paste(channels[i], 'vs.', channels[j]), pch = '.')
  }
}

par(mfrow = c(5,5), mar=c(2,2,2,2))
for (i in 1:dim(meanIntensity)[2]){
  for (j in 1:dim(meanIntensity)[2]){
    plot(meanIntensity_singleField[, i], meanIntensity_singleField[,j], xlab = channels[i], ylab = channels[j], bins = 50,
           main = paste(channels[i], 'vs.', channels[j]), pch = '.')
  }
}

```

See to what extent we can see some basic clusters in UMAP and hierarchical clustering:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE}

# intensityTab = CreateSeuratObject(log(t(meanIntensity),2))
# 
# intensityTab = ScaleData(intensityTab)
# intensityTab = RunPCA(intensityTab, features = c('Nuclei...Intensity.Nucleus.Atto.490LS.Mean', 'Nuclei...Intensity.Nucleus.Atto.425.Mean',
#                          'Nuclei...Intensity.Nucleus.Alexa.488.Mean', 'Nuclei...Intensity.Nucleus.Alexa.568.Mean',
#                          'Nuclei...Intensity.Nucleus.Alexa.647.Mean'))
# intensityTab = RunUMAP(intensityTab, dims = 1:4)
# UMAPPlot(intensityTab)
# PCAPlot(intensityTab)
# featurePlots = list()
# intensityFeatures = c('Nuclei...Intensity.Nucleus.Atto.490LS.Mean', 'Nuclei...Intensity.Nucleus.Atto.425.Mean',
#                          'Nuclei...Intensity.Nucleus.Alexa.488.Mean', 'Nuclei...Intensity.Nucleus.Alexa.568.Mean',
#                          'Nuclei...Intensity.Nucleus.Alexa.647.Mean')
# for (i in 1:length(intensityFeatures)){
#   print(i)
#   featurePlots[[i]] = FeaturePlot(intensityTab, features = intensityFeatures[[i]])
# }
# p = cowplot::plot_grid(featurePlots[[1]], featurePlots[[2]], featurePlots[[3]], featurePlots[[4]], 
#                        featurePlots[[5]])


Heatmap(meanIntensity)

```


Try out flowClust for clustering data into cell types:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE}

require(flowClust)

varNames = c('Nuclei...Intensity.Nucleus.Atto.490LS.Mean', 'Nuclei...Intensity.Nucleus.Atto.425.Mean',
                          'Nuclei...Intensity.Nucleus.Alexa.488.Mean', 'Nuclei...Intensity.Nucleus.Alexa.568.Mean',
                         'Nuclei...Intensity.Nucleus.Alexa.647.Mean')

par(mfrow = c(5,5), mar = c(1,1,1,1))
for (i in 1:length(varNames)){
  for (j in 1:length(varNames)){
    if (i != j){
          res2 <-
    flowClust(
        meanIntensity_singleField,
        varNames = varNames[c(i,j)],
        K = 3,
        B = 100
        )
    ruleOutliers(res2) <- list(level = 1)
    plot(res2,
       data = meanIntensity_singleField,
       level = 0.99,
      z.cutoff = 0, subset = varNames[c(i,j)], ellipse = F, main = paste(channels[i], 'vs.', channels[j]))
    }else{
    plot(1:10,1:10)
  }
}
}
```

